\rm Weak compactness of probability measures

### 2018-07-04 manuscript-3.tex
Adding the conclusion.

\section*{Measures as dual spaces}%\todo{write a few lines}

@@

\begin{theo}\label{charact}
With the above notations the linear mapping
\begin{equation*}
I\colon ba(\mathcal A) \to L^0_b(\mathcal A)^\prime, \quad \mu\mapsto I\mu \coloneqq%F_\mu\coloneqq
 \left(f\mapsto\int f\mathrm{d}\mu\right)
\end{equation*}
is an algebraic isomorphism and we have \(\left\lVert I\mu \right\rVert_{L^0_b(\mathcal A)^\ast} = \left\lVert \mu \right\rVert_{BV}\). In particular \(\left\lVert \cdot \right\rVert_{BV}\) is a norm and \(ba(\mathcal A)\) is complete with respect to it.
%isometric isomorphism.
 Further the inverse \(I^{-1}\) is given by
\[F\mapsto \Big( A\mapsto F(\chi_A)\Big).\]
For a measure \(\mu\in ba(\mathcal A)\) the following three statements are equivalent:
%is countably additive if and only if one of the two following equivalent properties holds:
\begin{enumerate}
\item \(\mu\) is countably additive.
\item \(I\mu\) fulfills the monotone convergence theorem.
\item \(I\mu\) fulfills the dominated convergence theorem.
\end{enumerate}
\end{theo}
\begin{proof}
The previous theorem directly implies that \(I\) is a well defined linear contraction, i.e. %It is clear that \(I\) is linear, bijective and well defined and %further the estimate \eqref{e1} yields %Further we have
%\[\mu(f)\le \int \left\lvert f \right\rvert\mathrm{d}\!\left\lvert \mu \right\rvert \le \left\lVert f \right\rVert_{L^0_b(\mathcal A)} \left\lVert \mu \right\rVert_{ba(\mathcal A)}, \]
%and therefore
\(\left\lVert I\mu \right\rVert\le \left\lVert \mu \right\rVert\). To see that the mapping is an isometry we take a finite collection \(\mathcal E\) of disjoint measurable sets. Then we get
\[\left\lVert I\mu \right\rVert_{L^0_b(\mathcal A)^\prime} \ge \int\limits \sum\limits_{E\in\mathcal E} \operatorname{sign}(\mu(E)) \chi_E\mathrm d\mu = \sum\limits_{E\in\mathcal E} \left\lvert \mu(E) \right\rvert. \]
%By choosing \(f\coloneqq \chi_{A_+} - \chi_{A_-}\), where \(A_+\) and \(A_-\) are the sets out of the Hahn-Jordan decomposition of \(\mu\) we get
%\[\mu(f) = \mu(A_+) - \mu(A_-) = \left\lVert \mu \right\rVert\]
By taking the supremum over all those collections \(\mathcal E\) we get \(\left\lVert I\mu \right\rVert = \left\lVert \mu \right\rVert\). Further the bijectivity follows from the fact \(I\circ I^{-1} =\operatorname{id}_{L^0_b(\mathcal A)^\prime}, I^{-1}\circ I = \operatorname{id}_{ba(\mathcal A)}\).  % \(I\mu_F = F\) and \(\mu_\)

@@

\begin{theo}
Let \(\mu\in ba(\mathcal A)\) and let \( ba(\mu)\subseteq ba(\mathcal A)\) be the subspace of measure that are absolutely continuous wrt \(\mu\). Then we have 
\[ ba(\mu) \cong L^\infty(\mu)^\prime \]
isometrically in the canonical way.
\end{theo}
\begin{proof}
Let \([f]_\mu\) denote the \(\mu\) ‚a.e. equivalence‘ class of \(f\in L^0_b(\mathcal A)\), then 
\[\iota:L^0_b(\mathcal A) \to L^\infty(\mu), \quad f\mapsto[f]_\mu\]
is surjective and further we have
\begin{equation}\label{e3}
\left\lVert [f]_\mu \right\rVert_{L^\infty(\mathcal A)} = \inf_{g\in [f]_\mu} \left\lVert g \right\rVert_\infty .
\end{equation}
Therefore the dual operator
\[\iota^\prime\colon L^\infty(\mu)^\prime \to L^0_b(\mathcal A)^\prime \cong ba(\mathcal A)\]
is injective and because of \eqref{e3} it is isometric since it implies \(\iota(B_1) = B_1\) where \(B_1\) denotes the unit ball in the respective spaces. %Indeed we have
%\[\left\lvert \iota^\prime F(f) \right\rvert = \left\lvert F[f]_\mu \right\rvert \le \left\lVert F \right\rVert_{L^\infty(\mu)^\prime} \cdot \left\lVert [f]_\mu \right\rVert_{L^\infty(\mu)} \le \left\lVert F \right\rVert_{L^\infty(\mu)^\prime}\cdot\left\lVert f \right\rVert_\infty\]
%and with Hahn Banach we get
%\[\left\lVert F \right\rVert_{L^\infty(\mu)^\prime} = F[f]_{\mu} = \lim \iota^\prime Fg_n\]
%with \((g_n)\subseteq [f]_\mu\) with \(\left\lVert g_n \right\rVert\to 1\).
Further the closed range theorem implies
\[\operatorname{ran}(\iota^\prime) = \operatorname{ker}(\iota)^\perp = \left\{ F\in L^0_b(\mathcal A)^\prime\mid F(f) = 0 \text{ for all } %f \text{ s.t. }
 f = 0 \text{ \(\mu\) a.e.}\right\} \cong ba(\mu).\]
\end{proof}

@@

In probability theory it is common to consider the \emph{weak topology} on the Borel probability measures \(\mathcal P(\mathcal A)\) over some topological space \(X\) with Borel algebra \(\mathcal A\), which is the initial topology of the evaluations
\[f\mapsto \mu(f) = \int f\mathrm{d}\mu\quad \text{for } f\in\mathcal C_b( X).\]
%We will now try to address the 
Note that this corresponds exactly to the relative topology on the probability measures \(\mathcal P(\mathcal A)\subseteq \mathcal C_b(X)^\ast\)  of the weak-\(\ast\) topology, where \(\mathcal P(\mathcal A)\) is embedded canonically.
If \(X\) is a complete metric space, then it is well known that the convergence that arrises from this topology, which is called the \emph{weak convergence}, can be metrised by the dual bounded Lipschitz metric or the Lévy-Prokhorov metric.
%\[d_{L}(\mu,\nu) \coloneqq \sup \Big\{ \mu(f) - \nu(f) \mid f\in\mathcal C_b(X) \text{ is Lipschitz with constant at most } 1 \Big\}.  \]
However this result is frequently stated as the fact that the weak topology is metrised with by this metric. Against the background of the metrisability results of the weak topologies in Banach spaces this seems very surprising and in fact it turns out to be false. More precisely the weak topology is metrisable if and only if the space \(X\) is compact, which is the case if and only if \(\mathcal C_b(X)\) is separable. This agrees with the metrisability of the weak-\(\ast\) topology.
To prove this statement, we will need the following ingredrients.
\begin{enumerate}
\item The space of probability measures over a topological space is compact in the weak topology.
\item If a famility \(\left\{ \mu_i\right\}_{i\in I}\) of probability measures is relatively sequentially compact in the weak topology, then it is \emph{tight}, i.e. for every \(\varepsilon>0\) there is a compact set \(K\subseteq X\) such that \(\mu_i(K)>1-\varepsilon\) for all \(i\in I\).
\item If \(X\) is a compact metric space, then \(\mathcal C_b(X)\) is separable. Note that the converse statement is also true, but we will not need it here.% if and only if \(X\) is compact.
\end{enumerate}
Assuming we have all those statements, then we can conclude as follows. If \(X\) is compact, then \(\mathcal C_b(X)\) is seperable and thus the weak-\(\ast\) topology on \(\mathcal C_b(X)^\ast\) is metrisable. In this case also the relative topology, i.e. the weak topology on the probability measures is metrisable. If on the other hand the weak topology is metrisable, then we know because of (\emph{i}) that the probability measures are sequentially compact. In particular the family of Dirac measures \(\left\{ \delta_x\right\}_{x\in X}\) is tight and thus there is a compact set \(K\subseteq X\) such that
\[\delta_x(K)>\frac12, \quad \text{i.e. } x\in K.\]
Hence we obtain that \(X=K\) is compact. 
Since (\emph{ii}) is well established in the literature (c.f. Prokhorov’s theorem or Theorem 5.2 in \cite{billingsley2013convergence}) and (\emph{iii}) follows from the Stone-Weierstrass theorem, we will focus on the proof of (\emph{i}).

@@

\section*{Conclusion}

>>>>

\section*{Measures as dual spaces}%\todo{write a few lines}
Now we have set up all the preliminaries that we need to perceive \(ba(\mathcal A)\) as a dual space.

@@

\begin{theo}[Measures as a dual space]\label{charact}
With the above notations the linear mapping
\begin{equation*}
I\colon ba(\mathcal A) \to L^0_b(\mathcal A)^\prime, \quad \mu\mapsto I\mu \coloneqq%F_\mu\coloneqq
 \left(f\mapsto\int f\mathrm{d}\mu\right)
\end{equation*}
is an algebraic isomorphism and we have \(\left\lVert I\mu \right\rVert_{L^0_b(\mathcal A)^\ast} = \left\lVert \mu \right\rVert_{BV}\). In particular \(\left\lVert \cdot \right\rVert_{BV}\) is a norm and \(ba(\mathcal A)\) is complete with respect to it.
%isometric isomorphism.
 Further the inverse \(I^{-1}\) is given by
\[F\mapsto \Big( A\mapsto F(\chi_A)\Big).\]
For a measure \(\mu\in ba(\mathcal A)\) the following three statements are equivalent:
%is countably additive if and only if one of the two following equivalent properties holds:
\begin{enumerate}
\item \(\mu\) is countably additive.
\item \(I\mu\) fulfills the monotone convergence theorem.
\item \(I\mu\) fulfills the dominated convergence theorem.
\end{enumerate}
\end{theo}
\begin{proof}
The previous theorem directly implies that \(I\) is a well defined linear contraction, i.e. %It is clear that \(I\) is linear, bijective and well defined and %further the estimate \eqref{e1} yields %Further we have
%\[\mu(f)\le \int \left\lvert f \right\rvert\mathrm{d}\!\left\lvert \mu \right\rvert \le \left\lVert f \right\rVert_{L^0_b(\mathcal A)} \left\lVert \mu \right\rVert_{ba(\mathcal A)}, \]
%and therefore
\(\left\lVert I\mu \right\rVert\le \left\lVert \mu \right\rVert\). To see that the mapping is an isometry we take a finite collection \(\mathcal E\) of disjoint measurable sets. Then we get
\[\left\lVert I\mu \right\rVert_{L^0_b(\mathcal A)^\prime} \ge \int\limits \sum\limits_{E\in\mathcal E} \operatorname{sign}(\mu(E)) \chi_E\mathrm d\mu = \sum\limits_{E\in\mathcal E} \left\lvert \mu(E) \right\rvert. \]
%By choosing \(f\coloneqq \chi_{A_+} - \chi_{A_-}\), where \(A_+\) and \(A_-\) are the sets out of the Hahn-Jordan decomposition of \(\mu\) we get
%\[\mu(f) = \mu(A_+) - \mu(A_-) = \left\lVert \mu \right\rVert\]
By taking the supremum over all those collections \(\mathcal E\) we get \(\left\lVert I\mu \right\rVert = \left\lVert \mu \right\rVert\). Further the bijectivity follows from the fact \(I\circ I^{-1} =\operatorname{id}_{L^0_b(\mathcal A)^\prime}, I^{-1}\circ I = \operatorname{id}_{ba(\mathcal A)}\).  % \(I\mu_F = F\) and \(\mu_\)

@@

\begin{theo}[Dual of \(L^\infty(\mu)\)]
Let \(\mu\in ba(\mathcal A)\) and let \( ba(\mu)\subseteq ba(\mathcal A)\) be the subspace of measure that are absolutely continuous wrt \(\mu\). Then we have 
\[ ba(\mu) \cong L^\infty(\mu)^\prime \]
isometrically in the canonical way.
\end{theo}
\begin{proof}
Let \([f]_\mu\) denote the \(\mu\) ‚a.e. equivalence‘ class of \(f\in L^0_b(\mathcal A)\), then 
\[\iota:L^0_b(\mathcal A) \to L^\infty(\mu), \quad f\mapsto[f]_\mu\]
is surjective and further we have
\begin{equation}\label{e3}
\left\lVert [f]_\mu \right\rVert_{L^\infty(\mathcal A)} = \inf_{g\in [f]_\mu} \left\lVert g \right\rVert_\infty .
\end{equation}
Therefore the dual operator
\[\iota^\prime\colon L^\infty(\mu)^\prime \to L^0_b(\mathcal A)^\prime \cong ba(\mathcal A)\]
is injective and because of \eqref{e3} it is isometric since it implies \(\iota(B_1) = B_1\) where \(B_1\) denotes the unit ball in the respective spaces. %Indeed we have
%\[\left\lvert \iota^\prime F(f) \right\rvert = \left\lvert F[f]_\mu \right\rvert \le \left\lVert F \right\rVert_{L^\infty(\mu)^\prime} \cdot \left\lVert [f]_\mu \right\rVert_{L^\infty(\mu)} \le \left\lVert F \right\rVert_{L^\infty(\mu)^\prime}\cdot\left\lVert f \right\rVert_\infty\]
%and with Hahn Banach we get
%\[\left\lVert F \right\rVert_{L^\infty(\mu)^\prime} = F[f]_{\mu} = \lim \iota^\prime Fg_n\]
%with \((g_n)\subseteq [f]_\mu\) with \(\left\lVert g_n \right\rVert\to 1\).
Further the closed range theorem implies
\[\operatorname{ran}(\iota^\prime) = \operatorname{ker}(\iota)^\perp = \left\{ F\in L^0_b(\mathcal A)^\prime\mid F(f) = 0 \text{ for all } %f \text{ s.t. }
 f = 0 \text{ \(\mu\) a.e.}\right\} \cong ba(\mu).\]
\end{proof}

@@

In probability theory it is common to consider the \emph{weak topology} on the Borel probability measures \(\mathcal P(\mathcal A)\) over some topological space \(X\) with Borel algebra \(\mathcal A\), which is the initial topology of the evaluations
\[f\mapsto \mu(f) = \int f\mathrm{d}\mu\quad \text{for } f\in\mathcal C_b( X).\]
%We will now try to address the 
Note that this corresponds exactly to the relative topology on the probability measures \(\mathcal P(\mathcal A)\subseteq \mathcal C_b(X)^\ast\)  of the weak-\(\ast\) topology, where \(\mathcal P(\mathcal A)\) is embedded canonically.
If \(X\) is a complete metric space, then it is well known that the convergence that arrises from this topology, which is called the \emph{weak convergence}, can be metrised by the dual bounded Lipschitz metric or the Lévy-Prokhorov metric.
%\[d_{L}(\mu,\nu) \coloneqq \sup \Big\{ \mu(f) - \nu(f) \mid f\in\mathcal C_b(X) \text{ is Lipschitz with constant at most } 1 \Big\}.  \]
However this result is frequently stated as the fact that the weak topology is metrised with by this metric. Against the background of the metrisability results of the weak topologies in Banach spaces this seems very surprising and in fact it turns out to be false. More precisely the weak topology is metrisable if and only if the space \(X\) is compact, which is the case if and only if \(\mathcal C_b(X)\) is separable. This agrees with the metrisability of the weak-\(\ast\) topology.
To prove this statement, we will need the following ingredrients.
\begin{enumerate}
\item The space of probability measures over a topological space is compact in the weak topology.
\item If a famility \(\left\{ \mu_i\right\}_{i\in I}\) of probability measures is relatively sequentially compact in the weak topology, then it is \emph{tight}, i.e. for every \(\varepsilon>0\) there is a compact set \(K\subseteq X\) such that \(\mu_i(K)>1-\varepsilon\) for all \(i\in I\).
\item If \(X\) is a compact metric space, then \(\mathcal C_b(X)\) is separable. Note that the converse statement is also true, but we will not need it here.% if and only if \(X\) is compact.
\end{enumerate}
Assuming we have all those statements, then we can conclude as follows. If \(X\) is compact, then \(\mathcal C_b(X)\) is seperable and thus the weak-\(\ast\) topology on \(\mathcal C_b(X)^\ast\) is metrisable. In this case also the relative topology, i.e. the weak topology on the probability measures is metrisable. If on the other hand the weak topology is metrisable, then we know because of (\emph{i}) that the probability measures are sequentially compact. In particular the family of Dirac measures \(\left\{ \delta_x\right\}_{x\in X}\) is tight and thus there is a compact set \(K\subseteq X\) such that
\[\delta_x(K)>\frac12, \quad \text{i.e. } x\in K.\]
Hence we obtain that \(X=K\) is compact. 
Since (\emph{ii}) is well established in the literature (c.f. Prokhorov’s theorem or Theorem 5.2 in \cite{billingsley2013convergence}) and (\emph{iii}) follows from the Stone-Weierstrass theorem (c.f. \cite{rudin1991functional}), we will focus on the proof of (\emph{i}).

@@

\section*{Conclusion}
We have seen that the Hahn-Jordan decomposition also holds for only finitely additive signed measures. To the authors knowledge there is no reference so far for this in the literature. Further we developed a duality between a function space and the Banach space of signed measures and proved that countable additivity is equivalent to the dominated or monotone convergence theorem. This duality provides an extremely short and self contained introduction to signed measures that avoids the mostly tedious calculation that are usually involved in the proof of the norm property of the total variation and the completeness of the space of signed measures.

In the last section it was shown that the set of probability measures are compact with respect to the weak topology. Further the motivating question was answered, namely it was the goal to investigate whether the Prokhorov metric metrises the weak topology like frequently stated or only the weak convergence (c.f. \cite{billingsley2013convergence}).% but frequently stated that it would metrise the weak topology itself.
% the question that was the motivation of this work was answered. Namely we wanted  It is well known that the Prokhorov metric metrises the weak convergence (c.f. \cite{billingsley2013convergence}) but frequently stated that it would metrise the weak topology itself.
% This implies that the weak topology of probability measures over a complete and separable metric space is metrisable if and only the space is compact. This answers the question which %was the motivation of the presented work whether the weak topology or only the weak convergence is metrised by the Prokhorov metric (c.f. \cite{billingsley2013convergence}).
In fact the weak topology of probability measures over a complete and separable metric space is metrisable if and only the space is compact and thus only the weak convergence is metrised. However this is no contradiction but shows that the weak topology has no countable neighborhood basis and is therefore not uniquely specified by its notion of convergence\todo{cite}.
